Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra

### 2022-03-03 analysis/paper/001-paper-main.Rmd
Update manuscript 1, update formatting fo references.

    To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra. All the advantages of such models rely on their critical validation. Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major compounds of organic matter in general. The models can be used to understand large-scale SOM gradients and have been used in various studies.  

@@

    The major limitation we identified is that the original training data are not representative for a range of peat samples which causes both biased estimates and extrapolation uncertainty under the original models. In addition, the original models can in parctice produce unrealistic predictions (negative values or values $>100$ mass-%). Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid unrealistic predictions. Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.  

@@

  We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template. We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions.

@@

```{r packages, echo=FALSE, include=FALSE}
# packages
library(ggplot2)
library(ir)
library(irpeat)
library(caret)
library(magrittr)
library(dplyr)
library(rstanarm)
library(brms)
library(loo)
library(projpred)
library(bayesplot)
library(patchwork)
library(ggrepel)
library(scales)
library(bib2df)
library(kableExtra)
library(purrr)
library(stringr)
library(tidyr)
```

@@

Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS. The models have several advantages over existing approaches. The training data comprises several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018]. Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a]. For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020], most recently in @Verbeke.2022, and for future developments, e.g. by including additional data.   

@@

Out-of-sample model performance was estimated with PSIS-LOO ELPD as described above [@Vehtari.2017; @Vehtari.2019]. If a model has a larger PSIS-LOO ELPD than another model fitted on the same data, it has a larger (leave-one-out cross-validation) predictive accuracy, so models with larger PSIS-LOO ELPD are preferred [@Vehtari.2017]. To compare models, one typically computes the posterior distribution of the difference in PSIS-LOO ELPD between the model with the largest PSIS-LOO ELPD and the other models, such that one can evaluate the probability of a certain predictive advantage of the best model in comparison to the other models [@Vehtari.2017; @Vehtari.2019].  
To further validate and interpret the models using binned spectra, we visually identified the most important bins (largest absolute median coefficients $\ge0.2$) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.

@@

Is the training data used to compute the models representative for the spectral properties of peat and peat forming vegetation? To answer this question, we needed to compare the spectral properties of the training samples to those of the peat and vegetation samples [@Wadoux.2021]: If for a sample the value of a spectral variable included in a model exceeds the range of values for the same variable in the training data, the model extrapolates if applied on the sample [@Wadoux.2021]. If extrapolation occurs, it is unclear if the predictions are valid estimates for holocellulose and Klason lignin contents.  
For the original models, the prediction domain is simply the range of area normalized heights of the `carb` and `arom15arom16` peak, respectively [@Hodgkins.2018]. Therefore, we could directly compare the area normalized heights of the `carb` and `arom15arom16` peak of the peat and vegetation samples to the range of area normalized heights of the same peaks for the training data.  
For the improved models using binned spectra, the predictor variables represent a multivariate prediction domain. We therefore created plots with which we could compare the value ranges for each bin for the training data with the respective values in the peat and vegetation spectra. This allowed us to identify for which spectral variables extrapolation is an issue. Finally, by identifying the most important variables in the improved models, we could qualitatively summarize how large the risk of this extrapolation is.  
We did not additionally investigate the prediction domain for the models using all extracted peaks and troughs since these models had no predictive advantage in comparison to the models using binned spectra.

## Analyzing differences in predictions between the original models and the improved models 
We analyzed how predictions of the improved models (best model for each of the two approaches) differ from the original models. We found considerable differences, even within the spectral range of the training data and therefore analyzed what factors probably cause these differences.  
In a first step, we were interested if the original or improved models are biased. @Hodgkins.2018 provide no plot of residuals against predicted values or measured values against predicted values. Both plots are commonly used to detect any biases in a model fit [e.g. @Pineiro.2008]. We therefore created such plots for the original model and the improved model and compared indication of bias between both.  
In a second step, we compared how predictions of both models differ in practice. To this end, we compared predictions of both for the peat and vegetation data from @Hodgkins.2018, similarly to how we compared the Gaussian and beta regression models. Since we found considerable differences, we conducted a more targeted exploratory analysis to identify factors probably causing these differences.

@@

In the introduction we mentioned that the original @Hodgkins.2018 model for holocellulose excluded training samples with admixtures of clay minerals (old magazine samples) since these interfere with the `carb` peak. By including more variables into the model than the `carb` peak (as described in the previous section), we tried to compute models that can also describe the holocellulose content in the training data samples with high clay content. For this, we fitted the best models for holocellulose from each of the two approaches described in the previous section again (binned spectra: `r loo_comparison_hol$bin_width[[1]]` cm$^{-1}$ bin width; extracted peaks: priors with standard deviation =`r loo_comparison_hol$prior_scale[loo_comparison_hol$variables == "peaks"][[1]]`), but this time including the four old magazine samples previously left out. We interpreted the coefficients of these binned models similarly to the versions not fitted with mineral rich samples.

@@

Our analysis shows that both strategies to include more variables result on average in more accurate predictions. If uncertainties for the estimated predictive accuracy are considered, the advantage is clear only for the models for holocellulose using binned spectra (Table \@ref(tab:res-tab-loo-comparison-res)), but not for Klason lignin. Consequently, using the data available, we could improve the model for holocellulose, but our analysis remains inconclusive with respect to Klason lignin.  
Is one of the strategies to include more variables advantageous? The on average overall best models, according to their PSIS-LOO ELPD, were those using binned spectra with bin widths of `r loo_comparison_hol$bin_width[[1]]` and `r loo_comparison_kl$bin_width[[1]]` for holocellulose and Klason lignin, respectively (Table \@ref(tab:res-tab-loo-comparison-res)). For Klason lignin, comparatively large prediction uncertainties result in the practical equivalence of both strategies. For holocellulose, using binned spectra (except bin width = 20) is more clearly better than using the extracted peaks. Summarized, this indicates that using binned spectra results in an improved predictive accuracy over the second approach to use all extracted peaks for holocellulose, whereas our validation remains inconclusive with respect to Klason lignin.  
To interpret the improved models using binned spectra, we plotted the median coefficients for the best models using binned spectra (Fig. \@ref(fig:res-p-coefficients-fig03)). From this plot, we identified the bins with the largest absolute median coefficients ($\ge0.2$, marked with points and labeled). The reason for the (on average) improved performance is that the improved models can use information contained in the complete spectra which is not contained in the manually selected peaks for the original models:  
Similarly to the original model, a variable near the `carb` peak is important to predict the holocellulose content (bin at ~950 cm$^{-1}$) since it is related to cellulose [@Cocozza.2003; @Stuart.2004; @Artz.2008], but two peaks related to aromatic structures are also important (~1200 cm$^{-1}$: C-O stretching in phenols and aromatic ethers [@Stuart.2004; @Kubo.2005]. ~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients-fig03), first panel). This indicates that aromatics provide information on the holocellulose content in the training samples. A plausible explanation is that holocellulose and Klason lignin together are the major mass fractions in many OM types and both are controlled by the same processes, but often in different directions (decomposition, resource allocation trade-off during plant growth, selective removal during processing) [@Biester.2014; @Chen.2016] (supporting Fig. 4<!---\@ref(fig:res-sup-kl-hol)--->).   
For Klason lignin, the improved model does not use bins near to the `arom15` and `arom16` peaks, in contrast to the original model. Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1270 and ~1160 cm$^{-1}$) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients-fig03), last panel). We have no explanation why one coefficient is negative and the other positive. In section "What causes differences in predictions?", we provide a mechanistic interpretation for why the selected variables are important.

@@

\clearpage

@@

The training data set does not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig. \@ref(fig:res-p-prediction-domain-fig04) (a) and (b)). For the original models, the training data `carb` and `arom15arom16` peak heights cover the range of the vegetation samples. However, few peat samples have larger `arom15arom16` and many peat samples have smaller `carb` peak heights than covered by the training data. This suggests that the original models extrapolate to predict holocellulose contents in peat samples with small holocellulose contents and Klason lignin contents in peat samples with large Klason lignin contents.  
The same is true for the improved models using binned spectra (Fig. \@ref(fig:res-p-prediction-domain-fig04) (c) and (d)). Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set. This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see section "Can holocellulose contents also be predicted for samples with mineral interferences?" below).  
Under what conditions does extrapolation occur? Samples with small `carb` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Holocellulose (with minerals)") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, are outside the prediction domain for holocellulose (Fig. \@ref(fig:res-p-prediction-domain-fig04)). Since the `carb` peak is small and absorbance at the selected wavenumbers is typically large in decomposed samples [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018] (and the respective coefficients are negative), this indicates that for holocellulose, extrapolation occurs mainly for more decomposed peat. 
This is also true for Klason lignin since extrapolation occurs for samples with large `arom15arom16` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Klason lignin") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, all of which are larger for more decomposed peat [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018].  
Overall, this indicates that the prediction domain formed by the training data does not cover peat --- particularly decomposed peat --- and partly peat forming vegetation samples. This is probably also true for non-peat SOM. Therefore the models' prediction can represent extrapolations in practice.

```{r res-p-prediction-domain-fig04, echo=FALSE, out.width="40%", fig.height=7, fig.width=4, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra (c, d). The prediction domain is the range of values covered by the predictor variables. The original models use only one predictor variable and therefore the predicition domain can be shown as spread of the training samples across the x axis (standardized height of the \\texttt{carb} and \\texttt{arom15arom16} peak, respectively) of a histogram (a, b; blue bars). Extrapolation occurs if other samples (peat, vegetation) exceed this range. Counts are number of samples in the respective datasets. The improved models include bins across the complete spectra (c, d). Therefore, the range in standardized predictor variable values (y axis) covered by the training data for each bin is shown as orange shaded area for the dataset with a bin width of 50 (c) and 10 cm$^{-1}$ (d), respectively, as used by the best improved models for holocellulose and Klason lignin. Extrapolation occurs if other samples (blue shaded regions) exceed these areas. Dashed lines indicate the most important variables for the holocellulose (c) and Klason lignin model (d), respectively (compare with Fig. \\ref{fig:res-p-coefficients-fig03})'}
p_prediction_domain
```

@@

In comparison to the training data, spectra for peat typically have larger absorption values at around 1200 cm$^{-1}$ and between (~1200 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig. 7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).  
These spectral differences are directly related to the differences in predicted values of the original and improved model (supplementary Fig. 8 <!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights. Similarly, the larger the trough at 1200 cm$^{-1}$ or the absorbance at 1550 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra. This indicates that both the delineation of the `carb` peak and predictions using binned spectra are potentially sensitive to these differences and that this causes the observed differences in the predictions even in the range of the training data where both models yield similar predictions for the training samples.  
What causes this sensitivity? Peak heights --- for the original model --- or spectral variables (bins) --- for the improved models using binned spectra --- are normalized by the sum of absorbance values in the complete spectra [@Hodgkins.2018]. This is necessary to make spectra comparable, but also renders inferences sensitive to absorbance of other peaks not directly considered or with low influence in the original and improved models, respectively. Thus, if the OH-peak is smaller, the normalized height of the `carb` peak is larger than for the same spectrum with a larger OH-peak. Likewise, for the improved model using binned spectra a smaller OH-peak results in larger absorbances around ~1200 to 1600 cm$^{-1}$ where the most influential variables are located (supplementary Fig. 7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)---> and 2<!---\@ref(fig:res-sup-spectra-sim-p))--->).  
Why do peat samples have such different spectral properties? We hypothesize the following mechanistic explanation for the differences: Decomposition of peat results in distinct changes in the absorbance at specific wavenumbers, e.g. the "OH-peak" and the fingerprint region [e.g. @Cocozza.2003]. For example, decomposition of phenols and disruption of tissue structures with hydrogen bonds result in a smaller absorbance around the OH-peak [@Kubo.2005; @Schellekens.2015]. The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (middle row in supplementary Fig. 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->). For decomposed OM, the same `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data.    
A consequence therefore is that it is questionable if the models (both original and improved) can be applied to peat and other SOM samples in general.

@@

We suggest that it is this bias which causes the difference in predictions of the original and improved models. Conditional on the relative contribution of the `arom15` peak to `arom15arom16`, peat samples have smaller and larger `arom15arom16` values than the training data (supplementary Fig. 10<!---\@ref(fig:res-p-pred-diff-kl-histogram1)--->). If one extrapolates the biased predictions of the original model for the training data classes, this means larger under- and overestimation for the peat samples than for the training data using the original model (Fig. \@ref(fig:res-p-pred-diff-train1-p2-fig06)). This is supported by measured peat Klason lignin contents: Peat from temperate peatlands was reported to have Klason lignin contents as high as ~70 mass-% and ~57 mass-% on average [@Hayes.2015]. Even though this was peat from extraction sites which probably is more decomposed [@Hayes.2015], this indicates that peat Klason lignin contents can be much larger, also in temperate regions, than predictions of the original model for wood rich tropical peat [@Hodgkins.2018]. In contrast, predictions under the improved model most probably reflect more correct Klason lignin estimates because the bias is much smaller and predicted values are larger for deeper peat samples. Therefore, this bias very likely is the reason why predictions of the original model differ from that of the improved model. Moreover, our results indicate that the improved model makes more correct predictions.  

@@

```{r references-preparation, echo=FALSE, eval=FALSE}
# get bibkeys
bibkeys <- bib2df::bib2df("references.bib")

# get texts to search for bibkey
si_text <- 
  readLines("008-paper-supplementary.Rmd") %>%
  paste(collapse = " ")

main_text <- 
  readLines("001-paper-main.Rmd") %>%
  paste(collapse = " ")

bibkeys <-
  bibkeys %>%
  dplyr::mutate(
    refs_only_si = 
      purrr::map_lgl(BIBTEXKEY, function(x) {
        stringr::str_detect(si_text, x) & !stringr::str_detect(main_text, x)
      })
  )

\nocite{`r paste(bibkeys$BIBTEXKEY[!bibkeys$refs_only_si], collapse = ", ")`}
\bibliography{references}

\nocitesi{`r paste(bibkeys$BIBTEXKEY[bibkeys$refs_only_si], collapse = ", ")`}
\bibliographystylesi{apacite}
\bibliographysi{references}
```

>>>>

    To understand global soil organic matter (SOM) chemistry and its dynamics, we need tools to efficiently quantify SOM properties, for example prediction models using mid infrared spectra. The advantages of such models rely on their validity and accuracy. Recently, \cite{Hodgkins.2018} developed models to predict organic matter holocellulose and Klason lignin contents, two indicators of SOM stability and major fractions of organic matter in general. The models can be used to understand large-scale SOM gradients and have been used in various studies.  

@@

    The major limitation we identified is that the original training data are not representative for a range of peat samples which causes both biased estimates and extrapolation uncertainty under the original models. In addition, the original models can in practice produce unrealistic predictions (negative values or values $>100$ mass-%). Our improved models partly reduce the observed bias, have a better predictive performance for the training data, and avoid unrealistic predictions. Finally, we provide a proof-of-concept that holocellulose contents can also be predicted for mineral-rich samples.  

@@

  We like to thank @Hodgkins.2018 and @DelaCruz.2016 for provision of data and code to reproduce and modify their analyses, and @Allaire.2020 for provision of an Rmarkdown template. We thank Suzanne Hodgkins and Jeff Chanton for constructive discussions which helped to improve a first version of this manuscript. We acknowledge support from the Open Access Publication Fund of the University of Münster.

@@

```{r packages, echo=FALSE, include=FALSE}
# packages
library(ggplot2)
library(ir)
library(irpeat)
library(caret)
library(magrittr)
library(dplyr)
library(rstanarm)
library(brms)
library(loo)
library(bayesplot)
library(patchwork)
library(ggrepel)
library(scales)
library(kableExtra)
library(purrr)
library(stringr)
library(tidyr)
```

@@

Recently, @Hodgkins.2018 developed models to predict Klason lignin and holocellulose contents from attenuated total reflectance MIRS. The models have several advantages over existing approaches. The training data comprises several different OM types, such as paper products (cardboard, office paper, magazines, newspaper), leaves from diverse species (trees and graminoids), and wood samples [@DelaCruz.2016; @Hodgkins.2018]. Moreover, both raw data and model code are publicly available [@Hodgkins.2018; @Teickner.2020a]. For this reason, the models are particularly suitable for application in other studies [@Teickner.2019; @Moore.2019; @Harris.2020; @Cong.2020; @Cong.2022], most recently in @Verbeke.2022, and for future developments, e.g. by including additional data.   

@@

Out-of-sample model performance was estimated with PSIS-LOO ELPD as described above [@Vehtari.2017; @Vehtari.2019]. If a model has a larger PSIS-LOO ELPD than another model fitted on the same data, it has a larger (leave-one-out cross-validation) predictive accuracy, so models with larger PSIS-LOO ELPD are preferred [@Vehtari.2017]. To compare models, one typically computes the posterior distribution of the difference in PSIS-LOO ELPD between the model with the largest PSIS-LOO ELPD and the other models, such that one can evaluate the probability of a certain predictive advantage of the best model in comparison to the other models [@Vehtari.2017; @Vehtari.2019].  
To further validate and interpret the models using binned spectra, we visually identified the most important bins in the models (largest absolute median coefficients $\ge0.2$) with bin a width of 20 cm$^{-1}$ (which are among the models with the best predictive accuracy; see below) and linked these to molecular structures and causal relations that most probably result in the correlation with holocellulose and Klason lignin, respectively.  
Based on the model validation, we defined the following set of models used in the subsequent analyses: "Best all peaks" denotes the models for holocellulose and Klason lignin with best average predictive accuracy of approach 1 described above. "Best binned spectra" denotes the models for holocellulose and Klason lignin with nearly the best average predictive accuracy of approach 2 described above. With "nearly the best predictive accuracy" we mean here that we used the models using binned spectra with a bin width of 20 cm$^{-1}$. This was the model with best average predictive accuracy for Klason lignin, but not holocellulose (Sect. 3.2). However, the model for holocellulose had a predictive performance very similar to the on average better model using a bin width of 10 cm$^{-1}$ (Sect. 3.2), and to keep the following analyses compact and easier to follow, we decided to use a bin width of 20 cm$^{-1}$ in all cases. These models are the models used during the subsequent analyses.

@@

Is the training data used to compute the models representative for the spectral properties of peat and peat forming vegetation? To answer this question, we needed to compare the spectral properties of the training samples to those of the peat and vegetation samples [@Wadoux.2021]: If for a sample the value of a spectral variable included in a model exceeds the range of values for the same variable in the training data, the model extrapolates if applied on the sample [@Wadoux.2021]. If extrapolation occurs, it is unclear if the predictions are valid estimates for holocellulose and Klason lignin contents.  
For the original models, the prediction domain is simply the range of area normalized heights of the `carb` and `arom15arom16` peak, respectively [@Hodgkins.2018]. Therefore, we could directly compare the area normalized heights of the `carb` and `arom15arom16` peak of the peat and vegetation samples to the range of area normalized heights of the same peaks for the training data.  
For the improved models using binned spectra, the predictor variables represent a multivariate prediction domain. We therefore created plots with which we could compare the value ranges for each bin for the training data with the respective values in the peat and vegetation spectra. This allowed us to identify for which spectral variables extrapolation is an issue. Finally, by identifying the most important variables in the improved models, we could qualitatively summarize how large the risk of this extrapolation is.  
Since models with a bin width of 20 cm$^{-1}$ were among the best models, we performed this analysis only for these models ("Best binned spectra"). We did not additionally investigate the prediction domain for the models using all extracted peaks and troughs ("Best all peaks") since these models had no predictive advantage in comparison to the models using binned spectra (Sect. 3.2).

## Analyzing differences in predictions between the original models and the improved models 
We analyzed how predictions of the improved models ("Best all peaks" and "Best binned spectra" defined in Sect. 2.2) differ from the original models. We found considerable differences, even within the spectral range of the training data, and therefore analyzed what factors probably cause these differences.  
In a first step, we were interested if the original or improved models are biased. @Hodgkins.2018 provide no plot of residuals against predicted values or measured values against predicted values. Both plots are commonly used to detect any biases in a model fit [e.g. @Pineiro.2008]. We therefore created such plots for the original model and the improved model and compared indication of bias between both.  
In a second step, we compared how predictions of the improved models differ in practice from the original models. To this end, we compared the models' predictions for the peat and vegetation data from @Hodgkins.2018, similarly to how we compared the Gaussian and beta regression models. Since we found considerable differences, we conducted a more targeted exploratory analysis to identify factors probably causing these differences.

@@

In the introduction we mentioned that the original @Hodgkins.2018 model for holocellulose excluded training samples with admixtures of clay minerals (old magazine samples) since these interfere with the `carb` peak. By including more variables into the model than the `carb` peak (as described in the previous section), we tried to compute models that can also describe the holocellulose content in the training data samples with high clay content. For this, we fitted the best models for holocellulose (the models in "Best all peaks" and "Best binned spectra" defined in Sect. 2.2) from each of the two approaches described in the previous section again, but this time including the four old magazine samples previously left out. We interpreted the coefficients of these binned models similarly to the versions not fitted with mineral rich samples (see Sect. 2.2).

@@

Our analysis shows that both strategies to include more variables result in on average more accurate predictions (Table \@ref(tab:res-tab-loo-comparison-res)). This can be concluded based on the average estimated ELPD values (larger values indicate a better estimated predictive accuracy). Consequently, using the data available, we could improve both the model for holocellulose and Klason lignin.  
Is one of the strategies to include more variables advantageous? The models with the largest predictive accuracy are models with small to moderate bin width (Table \@ref(tab:res-tab-loo-comparison-res)). These are better than the respective models using all extracted peaks: in addition to the ELPD, this can be derived from the standard errors of the ELPD differences relative to the best models ($2 \cdot \Delta \text{SE}$ is approximately the 95% confidence interval). In addition, the remaining binned models generally have a better average predictive accuracy than the models using all extracted peaks, except for the model with bin width = 100 cm$^{-1}$ for holocellulose. Summarized, this indicates that --- if a sufficiently high bin resolution (sufficiently small bin width) is chosen --- using binned spectra results in an improved predictive accuracy over the second approach to use all extracted peaks for both holocellulose and Klason lignin.  
To interpret the improved models using binned spectra, we plotted the median coefficients for the best models using binned spectra (Fig. \@ref(fig:res-p-coefficients-fig03)). From this plot, we identified the bins with the largest absolute median coefficients ($\ge0.2$, marked with points and labeled). The reason for the (on average) improved performance is that the improved models can use information contained in the complete spectra which is not contained in the manually selected peaks for the original models:  
Similarly to the original model, a variable near the `carb` peak is important to predict the holocellulose content (bin at ~970 cm$^{-1}$) since it is related to cellulose [@Cocozza.2003; @Stuart.2004; @Artz.2008], but one peak related to aromatic structures which is not extracted with the approach from @Hodgkins.2018 is also important (~1250 cm$^{-1}$: aromatic in plane C-H bending) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients-fig03), first panel). This indicates that aromatics provide information on the holocellulose content in the training samples and that extracting only selected peaks may omit useful information. A plausible explanation for the importance of this variable is that holocellulose and Klason lignin together are the major mass fractions in many OM types and both are controlled by the same processes, but often in different directions (decomposition, resource allocation trade-off during plant growth, selective removal during processing) [@Biester.2014; @Chen.2016] (supporting Fig. 4<!---\@ref(fig:res-sup-kl-hol)--->).   
For Klason lignin, the improved model does not use bins near to the `arom15` and `arom16` peaks, in contrast to the original model. Instead bins with large absolute coefficients are located in the fingerprint region and probably related to aromatic in plane C-H bending (~1150 and ~1270 cm$^{-1}$) [@Stuart.2004] (Fig. \@ref(fig:res-p-coefficients-fig03), last panel). We have no explanation why one coefficient is negative and the other positive. In Sect. 3.5<!--"What causes differences in predictions?"-->, we provide a mechanistic interpretation for why the selected variables are important.  
These selected bins are not represented in the extracted peaks. Consequently, models using all extracted peaks cannot use this information and tend to underfit. Similarly, binning with a broad bin width may result in a too coarse spectral resolution, as indicated by the relatively weak predictive accuracy of the model for holocellulose using binned spectra with a bin width of 100 cm$^{-1}$ (Table \@ref(tab:res-tab-loo-comparison-res)).

@@

@@

The training data set does not cover the range of the spectral variables relevant for predicting peat and peat forming vegetation holocellulose contents (Fig. \@ref(fig:res-p-prediction-domain-fig04) (a) and (b)). For the original models, the training data `carb` and `arom15arom16` peak heights cover the range of the vegetation samples. However, few peat samples have larger `arom15arom16` and many peat samples have smaller `carb` peak heights than covered by the training data. This suggests that the original models extrapolate to predict holocellulose contents in peat samples with small holocellulose contents and Klason lignin contents in peat samples with large Klason lignin contents.  
The same is true for the improved models using binned spectra (Fig. \@ref(fig:res-p-prediction-domain-fig04) (c)). Even though it is not possible to establish a direct relation between extrapolation for individual spectral variables and predicted contents, it is visible that for several bins, peat samples --- and to a lesser extent --- vegetation samples have mostly larger standardized predictor values than covered by the training data set. This is also true for bins with the largest absolute median coefficients (including the model for holocellulose fitted with mineral-rich samples; see Sect. 3.6 <!--"Can holocellulose contents also be predicted for samples with mineral interferences?"--> below).  
Under what conditions does extrapolation occur? Samples with small `carb` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Holocellulose (with minerals)") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, are outside the prediction domain for holocellulose (Fig. \@ref(fig:res-p-prediction-domain-fig04)). Since the `carb` peak is small and absorbance at the selected wavenumbers is typically large in decomposed samples [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018] (and the respective coefficients are negative), this indicates that for holocellulose, extrapolation occurs mainly for more decomposed peat. 
This is also true for Klason lignin since extrapolation occurs for samples with large `arom15arom16` peak, or for the binned spectra, large absorbance at ~`r m_pars %>% dplyr::filter(abs(median) >= 0.02 & variable_y == "Klason lignin") %>% dplyr::pull(wn) %>% sort() %>% knitr::combine_words()` cm$^{-1}$, all of which are larger for more decomposed peat [@Cocozza.2003; @Tfaily.2014; @Hodgkins.2018].  
Overall, this indicates that the prediction domain formed by the training data does not cover peat --- particularly decomposed peat --- and partly peat forming vegetation samples. This is probably also true for non-peat SOM. Therefore the models' predictions can represent extrapolations in practice; the training data is not in general representative for peat and peat-forming vegetation.

```{r res-p-prediction-domain-fig04, echo=FALSE, out.width="100%", fig.height=3, fig.width=9, fig.cap='Prediction domain for the original models (a, b), and the improved models using binned spectra with a bind width of 20 cm$^{-1}$ (c). The prediction domain is the range of values covered by the predictor variables. The original models use only one predictor variable and therefore the predicition domain can be shown as spread of the training samples across the x axis (standardized height of the \\texttt{carb} and \\texttt{arom15arom16} peak, respectively) of a histogram (a, b; blue bars). Extrapolation occurs if other samples (peat, vegetation) exceed this range. Counts are number of samples in the respective datasets. The improved models include bins across the complete spectra (c). Therefore, the range in standardized predictor variable values (y axis) covered by the training data for each bin is shown as orange shaded area for the dataset with a bin width of 20 cm$^{-1}$, as used by the (nearly) best improved models for holocellulose and Klason lignin. Extrapolation occurs if other samples (blue shaded regions) exceed these areas. Dashed lines indicate the most important variables for the holocellulose and Klason lignin model (compare with Fig. \\ref{fig:res-p-coefficients-fig03})'}
p_prediction_domain
```

@@

In comparison to the training data, spectra for peat typically have larger absorption values at around 1250 cm$^{-1}$ and between (~1250 to 1500 cm$^{-1}$), but a smaller absorption in the region of the "OH-peak" (~3400 cm$^{-1}$) (supplementary Fig. 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->).  
These spectral differences are directly related to the differences in predicted values of the original and improved model (supplementary Fig. 8<!---\@ref(fig:res-p-pred-diff-hol-intensity2)--->): The smaller the OH-peak is, the larger are predicted values of the original model in comparison to the improved model using binned spectra, especially for larger `carb` peak heights. Similarly, the larger the trough at 1250 cm$^{-1}$ or the absorbance at 1590 cm$^{-1}$ is, the larger are predictions of the original model in comparison to the improved model using binned spectra. This indicates that both the delineation of the `carb` peak and predictions using binned spectra are potentially sensitive to these differences and that this causes the observed differences in the predictions even in the range of the training data where both models yield similar predictions for the training samples.  
What causes this sensitivity? Peak heights --- for the original model --- or spectral variables (bins) --- for the improved models using binned spectra --- are normalized by the sum of absorbance values in the complete spectra [@Hodgkins.2018]. This is necessary to make spectra comparable, but also renders inferences sensitive to absorbance of other peaks not directly considered or with low influence in the original and improved models, respectively. Thus, if the OH-peak is smaller, the normalized height of the `carb` peak is larger than for the same spectrum with a larger OH-peak. Likewise, for the improved model using binned spectra a smaller OH-peak results in larger absorbances around ~1200 to 1600 cm$^{-1}$ where the most influential variables are located (supplementary Fig. 7 <!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)---> and 2<!---\@ref(fig:res-sup-spectra-sim-p))--->).  
Why do peat samples have such different spectral properties? We hypothesize the following mechanistic explanation for the differences: Decomposition of peat results in distinct changes in the absorbance at specific wavenumbers, e.g. the "OH-peak" and the fingerprint region [e.g. @Cocozza.2003]. For example, decomposition of phenols and disruption of tissue structures with hydrogen bonds result in a smaller absorbance around the OH-peak [@Kubo.2005; @Schellekens.2015]. The training data contain only undecomposed or industrially processed OM [@DelaCruz.2016; @Hodgkins.2018] which do not reflect such decomposition changes and therefore generally have a larger OH peak (middle row in supplementary Fig. 7<!---\@ref(fig:res-p-pred-diff-hol-spectra-diff2)--->). For decomposed OM, the same `carb` peak height or absorbance at specific wavenumbers therefore represents different holocellulose contents than for the training data.    
A consequence therefore is that it is questionable if the models (both original and improved) can be applied to peat and other SOM samples in general.

@@

We suggest that it is this bias which causes most of the difference in predictions of the original and improved models. Conditional on the relative contribution of the `arom15` peak to `arom15arom16`, peat samples have smaller and larger `arom15arom16` values than the training data (supplementary Fig. 10<!---\@ref(fig:res-p-pred-diff-kl-histogram1)--->). If one extrapolates the biased predictions of the original model for the training data classes, this means larger under- and overestimation for the peat samples than for the training data using the original model (Fig. \@ref(fig:res-p-pred-diff-train1-p2-fig06)). This is supported by measured peat Klason lignin contents: Peat from temperate peatlands was reported to have Klason lignin contents as high as ~70 mass-% and ~57 mass-% on average [@Hayes.2015]. Even though this was peat from extraction sites which probably is more decomposed [@Hayes.2015], this indicates that peat Klason lignin contents can be much larger, also in temperate regions, than predictions of the original model for wood rich tropical peat [@Hodgkins.2018]. In contrast, predictions under the improved model most probably reflect more correct Klason lignin estimates because the bias is much smaller and predicted values are larger for deeper peat samples. Therefore, this bias very likely is the reason why predictions of the original model differ from that of the improved model. Moreover, our results indicate that the improved model makes more correct predictions.  

@@
